---
title: "EDA_Submission"
author: "Gaurav Surtani"
date: "2023-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the iris dataset

```{r}
data(iris)
```

# Check for Missing Values

```{r}
missing_values <- sum(is.na(iris))
cat("Missing Values:", missing_values, "\n")
```

# Top Variables in the Dataset

```{r}
head(iris)
```

# Description of the Data

```{r}
num_observations <- nrow(iris)
num_variables <- ncol(iris)
cat("Number of Rows:", num_observations, "\n")
cat("Number of Columns:", num_variables, "\n")

numeric_variables <- sapply(iris, is.numeric)
cat("Numeric Variables:", sum(numeric_variables), "\n")

categorical_variables <- sapply(iris, is.factor)
cat("Categorical Variables:", sum(categorical_variables), "\n")

str(iris)
```

# Summary Statistics

```{r}
summary(iris)
summary(iris[numeric_variables])
```

# Summary Statistics for Numeric Variables

```{r}
summary(iris[, numeric_variables])
```

# Compute Statistics by Species

```{r}
summary_by_species <- aggregate(Sepal.Length ~ Species, data = iris, summary)
print(summary_by_species)
```

# Visualizations

## Scatterplot of Sepal Length vs. Sepal Width by Species

```{r}
library(ggplot2)
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(x = "Sepal Length", y = "Sepal Width", title = "Scatterplot of Sepal Length vs. Sepal Width by Species", subtitle = "Shows scatter of Width to Length")
```

## Trend of Petal Length vs. Petal Width by Species

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_line() +
  labs(x = "Petal Length", y = "Petal Width", title = "Trend of Petal Length vs. Petal Width by Species")
```

## Distribution of Sepal Length by Species

```{r}
ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_histogram(binwidth = 1.1, alpha = 0.7) +
  labs(x = "Sepal Length", y = "Frequency", title = "Distribution of Sepal Length by Species")
```

## Boxplot of Sepal Length by Species

```{r}
ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot() +
  labs(x = "Species", y = "Sepal Length", title = "Boxplot of Sepal Length by Species")
```